---
title: "Homework 2"
author: "Laura Lynch"
data: 2019-09-26
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
```

# Problem 1
Read and clean Mr. Trashwheel:
```{r}
trash_data = 
  read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")
    ) %>% 
  janitor::clean_names(dat = .) %>% 
  drop_na(data = ., dumpster) %>% 
  mutate(dat = ., 
         round(sports_balls, 0),
        as.integer(sports_ball))
trash_data
```

Read and clean precipitation data:
```{r}
```

About this data:
* number of observations in both resulting datasets 
```{r}
```
* examples of key variables. For available data
```{r}
```
* total precipitation in 2018
```{r}
```
* median number of sports balls in a dumpster in 2017
```{r}
```

# Problem 2
Clean and tidy pols-month:
```{r}
```
Clean and tidy snp:
```{r}
```
Clean and tidy snp:
```{r}
```
Merge the three datasets:
```{r}
```
About this data:
* Write a short paragraph about these datasets. Explain briefly 
* what each dataset contained
* describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables)

# Problem 3
Load and tidy data:
```{r}
```
Make a table:
```{r}
```
Make a scatterplot:
```{r}
```