---
title: "Homework 2"
author: "Laura Lynch"
data: 2019-09-26
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
```

# Problem 1 
Read and clean Mr. Trashwheel:
```{r}
trash_data = 
  readxl::read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N52"
    ) %>% 
  janitor::clean_names(dat = .) %>% 
  drop_na(data = ., dumpster) %>% 
  mutate(.data = ., 
        sports_balls = round(0),
        sports_balls_int = as.integer(sports_balls))
```

Read and clean precipitation data:
```{r}
rain_2018 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
    sheet = "2018 Precipitation", skip = 1, col_names = TRUE) %>%
    janitor::clean_names() %>%
  drop_na(total, month) %>%
  mutate(year = 2018)
rain_2018

rain_2017 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
    sheet = "2017 Precipitation", skip = 1, col_names = TRUE) %>%
    janitor::clean_names() %>%
  drop_na(total, month) %>%
  mutate(year = 2017)
rain_2017

rain_tidy = 
  bind_rows(rain_2018, rain_2017) %>%
  mutate(month = as.numeric(month), month = month.name[month]) %>%
  janitor::clean_names() %>%
  select(year, month, total) 
rain_tidy

```
About this data:
* number of observations in both resulting datasets 
The Mr. Trashwheel dataset has collected XX observations about XX dumpsters.
The Precipitation data HAS MONTHLY PRECIPITATION MEASURES OVER THE COURSE OF 2 YEARS.

* examples of key variables. For available data
For Mr. Trashwheel
* median number of sports balls in a dumpster in 2017
For the precipitation data
* total precipitation in 2018



# Problem 2
Clean and tidy pols-month:
```{r}
pols_month = readr::read_csv("data/pols-month.csv") %>%
  separate(mon, into = c("year", "month", "day"), sep = "-", convert = TRUE) %>%
  mutate(month = month.abb[month]) %>%
  mutate(president = ifelse(prez_gop == 1, c("gop"), c("dem"))) %>%
  select(-prez_dem, -prez_gop, -day) %>%
  arrange(year, month)
pols_month
```
  mutate(month = as.numeric(month), month = factor(month.name[month])) %>%

Clean and tidy snp:
```{r}
snp = readr::read_csv("data/snp.csv") %>%
  separate(date, into = c("day", "month", "year"), sep = "/", convert = TRUE) %>%
  mutate(month = month.abb[month]) %>%
  select(-day) %>%
  select(year, month, close)
snp
```
Clean and tidy unemployment:
```{r}
unemployment = readr::read_csv("data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    cols = jan:dec, 
    names_to = "month", 
    values_to = "percent_unemploy") %>%
    select(year, month) 
unemployment
```
        mutate(month = as.numeric(month), month = factor(month.name[month])) %>%
    mutate(month = month.name(month)) %>%
month = month.abb(month), 
mutate (month = as.numeric(month)) %>%
  #NAs introduced by coersion, trying to put full month
    mutate (month =as.factor(month.name["month"]) %>%

Merge the three datasets:
```{r}
# merge snp into pols
merge_1 = 
  left_join(pols_month, snp, by = c("year", "month"))
merge_1
# merge unemployment into merge_1
full_merge = 
  left_join(merge_1, unemployment, c("year", "month"))
full_merge
```
About this data:
* Write a short paragraph about these datasets. Explain briefly 
* what each dataset contained
* describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables)

# Problem 3
Load and tidy data:
```{r}
baby_names = readr::read_csv("data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  select(year = year_of_birth, 
         name = childs_first_name, 
         everything()) %>%
  mutate(name = str_to_lower(name)) %>%
  mutate(ethnicity = recode(ethnicity, 
         "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
         "BLACK NON HISP" = "BLACK NON HISPANIC", 
         "WHITE NON HISP" = "WHITE NON HISPANIC" 
         )) %>%
  distinct(name, year, ethnicity, .keep_all = TRUE)
```


Olivia's Table:
```{r}
baby_names %>% 
  filter(name == "olivia") %>%
  select(-count, -gender) %>%
  group_by(ethnicity) %>%
  pivot_wider(
   names_from = "year", 
   values_from = "rank") %>%
  knitr::kable() 
```
Boy's Table:
```{r}
baby_names %>% 
  filter(gender == "MALE", rank == 1) %>%
  select(-rank, -gender, -count) %>%
  group_by(ethnicity) %>%
  pivot_wider(
   names_from = "year", 
   values_from = "name") %>%
   knitr::kable() 
```

Make a scatterplot:
```{r}
baby_names %>%
  filter(gender == "MALE", year == 2016, ethnicity == "WHITE NON HISPANIC") %>%
  ggplot(aes(x = rank, y = count)) + geom_point()
```
, gender == "WHITE NON HISPANIC"