---
title: "Homework 2"
author: "Laura Lynch"
data: 2019-09-26
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(dplyr)
```

# Problem 1 
Read and clean Mr. Trashwheel:
data/~$HealthyHarborWaterWheelTotals2018-7-28.xlsx

```{r, warning = FALSE}
trash_data = 
  readxl::read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N338",
    ) %>% 
  janitor::clean_names(dat = .) %>% 
  drop_na(data = ., dumpster) %>% 
  mutate( 
        sports_balls = as.integer(sports_balls, round, 0)
        )
```

Read and clean precipitation data:
```{r, warning = FALSE}
rain_2018 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
    sheet = "2018 Precipitation", skip = 1, col_names = TRUE) %>%
    janitor::clean_names() %>%
  drop_na(total, month) %>%
  mutate(year = 2018)

rain_2017 = 
  readxl::read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
    sheet = "2017 Precipitation", skip = 1, col_names = TRUE) %>%
    janitor::clean_names() %>%
  drop_na(total, month) %>%
  mutate(year = 2017)

rain_tidy = 
  bind_rows(rain_2018, rain_2017) %>%
  mutate(month = as.numeric(month), month = month.name[month]) %>%
  janitor::clean_names() %>%
  select(year, month, total) 
```
About the data:

 * The Mr. Trashwheel is collected from `r nrow(trash_data)` dumpsters
    * Measurements are collected for `r ncol(trash_data)` types of debris pulled from the ocean
        * For example the median number of sports balls removed was `r median(pull(trash_data, sports_balls))`
        * And a median of `r median(pull(trash_data, grocery_bags))` grocery bags
    * In total `r median(pull(trash_data, weight_tons))` tons of garbage was removed since 2014!

 * The 2017-2018 precipitation data has `r nrow(rain_tidy)` observations
    * The total precipitation for 2017 was `r sum(pull(rain_2017, total))` inches
    * The total precipitation for 2018 was `r sum(pull(rain_2018, total))` inches


# Problem 2
Clean and tidy pols-month:
```{r, warning = FALSE}
pols_month = readr::read_csv("data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-", convert = TRUE) %>%
  mutate(month = month.abb[month]) %>%
  mutate(president = ifelse(prez_gop == 1, c("gop"), c("dem"))) %>%
  select(-prez_dem, -prez_gop, -day) %>%
  arrange(year, month)
```
  

Clean and tidy snp:
```{r, warning = FALSE}
snp = readr::read_csv("data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("day", "month", "year"), sep = "/", convert = TRUE) %>%
  mutate(month = month.abb[month]) %>%
  select(-day) %>%
  select(year, month, close)
```
Clean and tidy unemployment:
```{r, warning = FALSE}
unemployment = readr::read_csv("data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    cols = jan:dec, 
    names_to = "month", 
    values_to = "percent_unemploy") %>%
    select(year, month) 
```

Merge the three datasets:
```{r}
# merge snp into pols
merge_1 = 
  right_join(pols_month, snp, by = c("year", "month"))

# merge unemployment into merge_1
full_merge = 
  left_join(merge_1, unemployment, c("year", "month"))
```

The"pols_month" dataset tracks the political affiliations of national leaders monthly including president, senators and governers for a total of `r nrow(pols_month)` months. The "snp" dataset tracks the closing values of the Standard & Poors stock from  `r min(pull(snp, year))` to `r max(pull(snp, year))`. Finally, the "unemployment" dataset tracks the percentage unemployment by month from `r min(pull(unemployment, year))` to `r max(pull(unemployment, year))`.

Once merged, these data represent `r nrow(full_merge)` observations and `r ncol(full_merge)` variables. The data collected range frorm `r min(pull(full_merge, year))` to `r max(pull(full_merge, year))`. It is possible to see how many democrats and republicans hold high office, including the political affiliation of the president, and a break down of S&P closing values by month. The mean S&P clsoing value was `r mean(pull(full_merge, close))`, the median S&P closing value was `r median(pull(full_merge, close))` and the standard deviation was `r sd(pull(full_merge, close))`.


# Problem 3
Load and tidy data:
```{r, warning = FALSE}
baby_names = readr::read_csv("data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  select(year = year_of_birth, 
         name = childs_first_name, 
         everything()) %>%
  mutate(name = str_to_lower(name)) %>%
  mutate(ethnicity = recode(ethnicity, 
         "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER",
         "BLACK NON HISP" = "BLACK NON HISPANIC", 
         "WHITE NON HISP" = "WHITE NON HISPANIC" 
         )) %>%
  distinct(name, year, ethnicity, .keep_all = TRUE)
```

Olivia's Table:
```{r}
baby_names %>% 
  filter(name == "olivia", gender == "FEMALE") %>%
  select(-count, -gender) %>%
  group_by(ethnicity) %>%
  pivot_wider(
   names_from = "year", 
   values_from = "rank") %>%
  knitr::kable() 
```


Boy's Table:
```{r}
baby_names %>% 
  filter(gender == "MALE", rank == 1) %>%
  select(-rank, -gender, -count) %>%
  group_by(ethnicity) %>%
  pivot_wider(
   names_from = "year", 
   values_from = "name") %>%
   knitr::kable() 
```

Make a scatterplot:
```{r}
baby_names %>%
  filter(gender == "MALE", year == 2016, ethnicity == "WHITE NON HISPANIC") %>%
  ggplot(aes(x = rank, y = count)) + geom_point() + 
  labs(
    title = "NYC Baby Names",
    x = "Popularity Ranking",
    y = "Number of Babies",
    caption = "Data from NYC OpenData, Popular Baby Names"
  )
```
